---
title: "1. Accessing Human Cell Atlas Data From Bioconductor"
author:
    Martin Morgan^[Roswell Park Comprehensive Cancer Center (RPCCC)],
    Maya McDaniel,
    Yubo Cheng^[RPCCC].
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{1. Accessing Human Cell Atlas Data From Bioconductor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Last modified: 20 July, 2022

## Overview

### Description

FIXME: Along with the topic of your workshop, include how students can expect
to spend their time. For the description may also include information
about what type of workshop it is (e.g. instructor-led live demo, lab,
lecture + lab, etc.). Instructors are strongly recommended to provide
completely worked examples for lab sessions, and a set of stand-alone
notes that can be read and understood outside of the workshop.

### _R_ / _Bioconductor_ packages used

The focus is on the [hca][] package. This package emphasize [dplyr][]
and 'tidy' approaches to working with data.frames. Files downloaded
from the HCA or CellXGene web sites can be imported into R /
Bioconductor as [SingleCellExperiment][] objects through the
[LoomExperiment][] and [zellkonverter][] packages.

[hca]: https://bioconductor.org/packages/hca
[cellxgenedp]: https://bioconductor.org/packages/cellxgenedp
[dplyr]: https://cran.r-project.org/package=dplyr
[SingleCellExperiment]: https://bioconductor.org/packages/SingleCellExperiment
[LoomExperiment]: https://bioconductor.org/packages/LoomExperiment
[zellkonverter]: https://bioconductor.org/packages/zellkonverter

### Time outline

| Activity                            | Time |
|-------------------------------------|------|
| Introduction to the HCA data portal |  5m  |
| HCA discovery in R / Bioconductor   | 10m  |

### Workshop goals and objectives

List "big picture" student-centered workshop goals and learning
objectives. Learning goals and objectives are related, but not the
same thing. These goals and objectives will help some people to decide
whether to attend the conference for training purposes, so please make
these as precise and accurate as possible.

*Learning goals* are high-level descriptions of what
participants will learn and be able to do after the workshop is
over. *Learning objectives*, on the other hand, describe in very
specific and measurable terms specific skills or knowledge
attained. The [Bloom's Taxonomy](#bloom) may be a useful framework
for defining and describing your goals and objectives, although there
are others.

### Learning goals

Some examples:

* describe how to...
* identify methods for...
* understand the difference between...

### Learning objectives

* analyze xyz data to produce...
* create xyz plots
* evaluate xyz data for artifacts

## Workshop

### Exploring the HCA and CellXGene web sites

Human Cell Atlas data portal

- https://data.humancellatlas.org/explore/projects

CellXGene

- https://cellxgene.cziscience.com/
- Easy visualization is pretty cool! 

### Working with HCA data

### Working with CellXGene data

### Importing .loom or .h5ad data into _R_ / _Bioconductor_

### Integration with _R_ / _Bioconductor_ work flows

## Acknowledgements

```{r}
sessionInfo()
```

